---
output: github_document
---


[![Project Status: WIP - Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](http://www.repostatus.org/badges/latest/wip.svg)](http://www.repostatus.org/#wip)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/epiflows)](https://cran.r-project.org/package=epiflows)
[![Travis-CI Build Status](https://travis-ci.org/reconhub/epiflows.svg?branch=master)](https://travis-ci.org/reconhub/epiflows)


# Welcome to the *epiflows* package!

`epiflows` is a package for predicting and visualising spread of infectious diseases based on flows between geographical locations, e.g., countries.
`epiflows` provides functions for calculating spread estimates, handling flow data, and visualization.

## Installing the package

To install the current stable, CRAN version of the package, type:
```{r install, eval = FALSE}
install.packages("epiflows")
```

To benefit from the latest features and bug fixes, install the development, *github* version of the package using:
```{r install2, eval = FALSE}
devtools::install_github("reconhub/epiflows")
```

Note that this requires the package *devtools* installed.


# What does it do?

The main features of the package include:

* `epiflows`: an S3 class for storing flow data, as well as country metadata
* `make_epiflows`: a constructor for `epiflows`
* `add_coordinates`: add latitude/longitude to the location data in an `epiflows` object using `ggmap::geocode()`
* `fn_number_cases_spread`: calculate estimates (point estimate and 95% CI) for disease spread from flow data
* `get_flow_data`: return flow data to and/or from specified location
* `get_location_data`: return metadata for specified location(s)
* `x[i]`: subset an `epiflows` object to location(s) *i*
* `plot`: plot flows from an `epiflows` object on a *leaflet* world map
* `print`: print summary for an `epiflows` object

# Resources

## Vignettes

An overview and examples of *epiflows* are provided in the vignettes:

...

## Getting help online

Bug reports and feature requests should be posted on *github* using the [*issue*](http://github.com/reconhub/epiflows/issues) system. All other questions should be posted on the **RECON forum**: <br>
[http://www.repidemicsconsortium.org/forum/](http://www.repidemicsconsortium.org/forum/)

Contributions are welcome via **pull requests**.

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

